---
title: "New Analysis"
author: "WAK"
date: "today"
format:
  html: 
    toc: true
    toc-location: left
editor: visual
---

```{r}
#| echo: false
#| include: false
#| warning: false

library(epiDisplay)
library(knitr)
library(pROC)
library(tidyverse)
library(readxl)
library(ggplot2)
library(gmodels)
library(ggbeeswarm)
library(dplyr)
library(ggpubr)
library(psych)
library(lme4)
library(Matrix)
library(lmerTest)
library(irr)
library(lmerTest)
library(emmeans)
library(doBy)       
library(gmodels)    
library(car)           
library(cowplot)       
library(gridGraphics)  
library(multcomp)
library (broom)
library(DT)
library(gt)
library(survival)
library(survminer)
```

```{r}
#| echo: false
#| include: false
#| warning: false

#import the data

treatment <- read_csv ("data/ENROLL FINAL 04062022.csv")

#create data set ("new") excluding all exclusions 

new <- treatment |> 
  filter (ExYN10 == 0)


wths <- read_csv ("data/Weight HS Final 04072022.csv")
wths$week <- as.factor(wths$week)

wths_noex <- wths |> 
  filter(ExYN10 == 0)


```

## New Analysis for Disbudding Y2

This analysis is a new one for the disbudding Y2 data collected during summer of 2021. We enrolled 560 goat kids in a disbudding trial to investigate the effect of multi-modal pain management vs. a positive control.

## Descriptive Table of Enrolled Goat Kids

```{r}
#| echo: false


table1 <- new |> 
  group_by(Treatment) |> 
  summarize(
    avgbwt = round(mean(bwtkg, na.rm = TRUE), 2),
    sdbwt = round(sd(bwtkg, na.rm = TRUE), 2), 
    avgdia = round(mean(DIAenroll, na.rm = TRUE), 2),
    sddia = round(sd(DIAenroll, na.rm = TRUE), 2),
    avgewt = round(mean(weight, na.rm = TRUE), 2), 
    sdewt = round(sd(weight, na.rm = TRUE), 2)
  ) |> 
  ungroup()

new |> 
  group_by(Breed) |> 
  summarize(
    avgbwt = round(mean(weight, na.rm = TRUE), 2),
    sdbwt = round(sd(weight, na.rm = TRUE), 2), 
    avgdia = round(mean(DIAenroll, na.rm = TRUE), 2),
    sddia = round(sd(DIAenroll, na.rm = TRUE), 2)) |> 
  ungroup()



  
gt(table1) 

new |> 
  group_by (Treatment, Sex) |> 
  count() |> 
  ungroup()

CrossTable(new$Treatment,  new$Sex, prop.r=FALSE, prop.c = FALSE, fisher=TRUE)


new |> 
  group_by(Treatment, Breed) |> 
  count() |> 
  ungroup()

CrossTable(new$Treatment,  new$Breed, prop.r=FALSE, prop.c = FALSE, fisher=TRUE)


t.test(new$weight~new$Treatment)
shapiro.test(new$weight)

new |> 
  ggplot (mapping = aes(x=weight))+
          geom_histogram()
        


t.test(new$DIAenroll~new$Treatment)
shapiro.test(new$DIAenroll)

new |> 
  ggplot (mapping = aes(x=DIAenroll))+
          geom_histogram()
        

  
  

```

## More enrollment and day of disbudding

```{r}
#| echo: false
#| warning: false

wths_noex |> 
  group_by(Treatment, week) |> 
  summarize(
    wt = round(mean(weight, na.rm = TRUE), 2),
    sdwt = round(sd(weight, na.rm = TRUE), 2), 
  ) |> 
  ungroup()

week0 <- wths_noex |> 
  filter (week == 0) 

t.test(week0$weight~week0$Treatment)


```

## Descriptive Stuff for Health Scores for first three weeks

```{r}
#| echo: false
#| warning: false

first3 |> 
  group_by(Treatment) |> 
  count(Mentation) 

first3 |> 
  group_by(Treatment) |> 
  count(bodyposture)

first3 |> 
  group_by(Treatment) |> 
  count(resp)

first3 |> 
  group_by(Treatment) |> 
  count(bellyfill)

first3 |> 
  group_by(Treatment) |> 
  count(fecal)

first3 |> 
  group_by(Treatment) |> 
  count(rectal_temp)

first3 |> 
  group_by(Treatment) |> 
  count(Total_HS)
  

```

## Filtering data to just first 3 weeks for weight and hs data

```{r}
#| echo: false
#| warning: false

first3 <- 
  wths_noex |> 
  filter (week %in% c("0", "1", "2", "3"))

#composite sick score by week
first3 |> 
  group_by(Treatment) |> 
  count(composite)

first3 |> 
  group_by(Treatment, week) |> 
  count(composite)

#model for first three weeks

model <- glmer(composite ~ Treatment +  Breed + (1 | Kidid) + (1|pen), family = "binomial", data = first3)
summary(model)
emmeans(model, specs = "Treatment", at = list(Treatment = c("Trt", "Con")), type="response")

fixed_effects <- fixef(model)
odds_ratios <- exp(fixed_effects)
cat("Odds Ratios:\n")
print(odds_ratios)
type3_p_values <- Anova(model, type="III", test="Chisq")
cat("\nType III p-values:\n")
print(type3_p_values)

#model for six weeks

wths_noex |> 
  group_by(Treatment) |> 
  count(composite)

model <- glmer(composite ~ Treatment +  Breed + (1 | Kidid) + (1|pen), family = "binomial", data = wths_noex)
summary(model)
emmeans(model, specs = "Treatment", at = list(Treatment = c("Trt", "Con")), type="response")

fixed_effects <- fixef(model)
odds_ratios <- exp(fixed_effects)
cat("Odds Ratios:\n")
print(odds_ratios)
type3_p_values <- Anova(model, type="III", test="Chisq")
cat("\nType III p-values:\n")
print(type3_p_values)






```

## Health Treatment Information

```{r}
#| echo: false
#| warning: false

new |> 
  group_by (Treatment) |> 
  count(trt0yn10)

new |> 
  group_by(Treatment) |> 
  count(trt1yn10)

new |> 
  group_by (Treatment) |> 
  count(Dead_YN10)


```

## Kaplan-Meyer Curves for all data (6weeks) for Treatment post disbudding and Dead post disbudding

```{r}
#| echo: false
#| warning: false


#Univariable model for Treatment Survival
km_fit <- survfit(Surv(time = new$DFDT1, event = new$trt1yn10) ~ Treatment, data = new)


summary(km_fit)
plot(km_fit, 
      xlab = "Time (days)", 
     ylab = "Survival Probability", 
     main = "Kaplan-Meier Survival Curve", 
     col = c("blue", "red"), 
     mark.time = FALSE)
legend("topright", legend = c("Group 1", "Group 2"), col = c("blue", "red"), lty = 1:2)
km_fit$median
print(km_fit, print.rmean = TRUE)

#wilcoxon test will give more weight to early events vs. later events

wilcoxon_test<-survdiff(Surv(time = new$DFDT1, event = new$trt1yn10) ~ Treatment, data = new, rho = 1)




#Univariable model for Regular Survival
km_fit <- survfit(Surv(time = new$DFDDie, event = new$trt1yn10) ~ Treatment, data = new)


summary(km_fit)
plot(km_fit, 
      xlab = "Time (days)", 
     ylab = "Survival Probability", 
     main = "Kaplan-Meier Survival Curve", 
     col = c("blue", "red"), 
     mark.time = FALSE)
legend("bottomleft", legend = c("Group 1", "Group 2"), col = c("blue", "red"), lty = 1:2)
km_fit$median
print(km_fit, print.rmean = TRUE)

#wilcoxon test will give more weight to early events vs. later events

wilcoxon_test<-survdiff(Surv(time = new$DFDDie, event = new$trt1yn10) ~ Treatment, data = new, rho = 1)

wilcoxon_test





#Univariable model for Diarrhea
km_fit <- survfit(Surv(time = new$DFDT1, event = new$Trt1diar) ~ Treatment, data = new)


summary(km_fit)
plot(km_fit, 
      xlab = "Time (days)", 
     ylab = "Survival Probability", 
     main = "Kaplan-Meier Survival Curve", 
     col = c("blue", "red"), 
     mark.time = FALSE)
legend("bottomleft", legend = c("Group 1", "Group 2"), col = c("blue", "red"), lty = 1:2)
km_fit$median
print(km_fit, print.rmean = TRUE)

#wilcoxon test will give more weight to early events vs. later events

wilcoxon_test<-survdiff(Surv(time = new$DFDT1, event = new$Trt1diar) ~ Treatment, data = new, rho = 1)

wilcoxon_test




#Univariable Models for Pneumonia Treatment 

km_fit <- survfit(Surv(time = new$DFDT1, event = new$Trt1pneu) ~ Treatment, data = new)


summary(km_fit)
plot(km_fit, 
      xlab = "Time (days)", 
     ylab = "Survival Probability", 
     main = "Kaplan-Meier Survival Curve", 
     col = c("blue", "red"), 
     mark.time = FALSE)
legend("bottomleft", legend = c("Group 1", "Group 2"), col = c("blue", "red"), lty = 1:2)
km_fit$median
print(km_fit, print.rmean = TRUE)

#wilcoxon test will give more weight to early events vs. later events

wilcoxon_test<-survdiff(Surv(time = new$DFDT1, event = new$Trt1pneu) ~ Treatment, data = new, rho = 1)

wilcoxon_test




```

## Playing around with KM formatting

```{r}

km_fit <- survfit(Surv(time = new$DFDT1, event = new$Trt1pneu) ~ Treatment, data = new)
ggsurvplot(km_fit,
           data = new,         # The dataset
           conf.int = TRUE,     # Add shaded confidence intervals
           conf.int.style = "ribbon",  # Style of the confidence intervals
           palette = c("blue", "red"), # Customize color palette
           xlab = "DFDT1",      # Custom X-axis label
           ylab = "Survival Probability",  # Custom Y-axis label
           legend.title = "Treatment",    # Custom legend title
           legend.labs = c("Con", "Trt"), # Custom legend labels
           pval = TRUE,         # Display p-value for log-rank test
           risk.table = TRUE)   # Optionally add risk table below plot
```
